---
title: "Lab 1 - Data visualization"
author: "Lexie Brzoska"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

## Load Packages

```{r}
#1 label: load-packages
#1 message: false

library(tidyverse)
library(viridis)
```

## Exercise 1

The histogram has a right skew and does have some outliers nearing the 100,000
population density.
```{r}
#2 label: Histogram for population density of counties

ggplot(midwest) +
  geom_histogram( aes( popdensity ), binwidth = 10000) +
  labs(
    x= "Pop Density",
    y = "Count",
    title = "Pop Density in each County",
  )

```


## Exercise 2

```{r}
#3 label: Scatterplot for the percentage of people with a college degree (percollege) versus percentage below poverty (percbelowpoverty)

ggplot(midwest, aes( x= percollege, y= percbelowpoverty, color= state ) ) +
  geom_point(show.legend= TRUE) +
labs(
    x= "Percent College Educated",
    y = "Percent Below Poverty",
    title = "Percollege vs. Percbelowpoverty")
  


```


## Exercise 3

In the plot from exercise 2, the scatter is more concentrated in between 10 and
20 % college educated. There is an outlier around the 50% college educated. 

## Exercise 4

I think I prefer this new graph from excericise 4. It seems clearer and easier
to understand the most important information. I bet it depends on the situation,
though. 
```{r}
#4 label: Ex-4, using facet_wrap

ggplot(midwest, aes( x= percollege, y= percbelowpoverty, color= state ) ) +
  geom_smooth( se= FALSE, show.legend= TRUE) +
labs(
    x= "Percent College Educated",
    y = "Percent Below Poverty",
    title = "Percollege vs. Percbelowpoverty")+
  facet_wrap('state')
  
```


## Exercise 5
From this graph I can see that IN and OH both have the smallest counties. However,
there isn't a drastic difference in the number between the states and sizes of their
counties. WI ends up having the largest county size. I know because their centralized
data is overall higher than the others. 

```{r}
#5 label: Ex-5, side-by-side box plots

library(ggplot2)
ggplot(midwest, aes( x= state, y= area, fill = state) ) +
  geom_boxplot( show.legend= TRUE, alpha=.5) +
labs(
    x= "State",
    y = "Area",
    title = "Area vs.County by State")
```


## Exercise 6

Looking at the plot, the percentage of counties in metro areas is highest in Indiana and Ohio. Additionally, pretty much all counties have fewer in metro areas. 
```{r}
#6 label: Ex-6, Segmented bar graph

midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No")) 
ggplot(midwest, 
       aes( x= state, fill = metro) ) +
  geom_bar( position = "fill") +
labs(
    x= "State",
    y = "% of Counties",
    title = "States vs. Counties in the Metro")
```


## Exercise 7
```{r}
#7 label: Ex-7, replicating scatterplot


ggplot(midwest,
       aes( x= percollege, y= popdensity, color = percbelowpoverty) ) +
  geom_point( size = 2, alpha=.5) +
 facet_wrap('state')+
labs(
    x= "% College Educated",
    y = "Population Density (person/unit area)",
    title = "Do people with college degrees tend to live in denser areas?", 
    color= "% below \npoverty line") +
  theme_minimal()
```

