---
title: "Lab 1 - Data visualization"
author: "Lexie Brzoska"
format: pdf
editor: visual
---

## Load Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(viridis)
```

## Exercise 1

The histogram has a right skew and does have some outliers nearing the 100,000
population density.
```{r}
#| label: Histogram for population density of counties

ggplot(midwest) +
  geom_histogram( aes( popdensity ), binwidth = 10000) +
  labs(
    x= "Pop Density",
    y = "Count",
    title = "Pop Density in each County",
  )

```


## Exercise 2

```{r}
#| label: Scatterplot for the percentage of people with a college degree (percollege) versus percentage below poverty (percbelowpoverty)

ggplot(midwest, aes( x= percollege, y= percbelowpoverty, color= state ) ) +
  geom_point(show.legend= TRUE) +
labs(
    x= "Percent College Educated",
    y = "Percent Below Poverty",
    title = "Percollege vs. Percbelowpoverty")
  


```


## Exercise 3

In the plot from exercise 2, the scatter is more concentrated in between 10 and
20 % college educated. There is an outlier around the 50% college educated. 

## Exercise 4

I think I prefer this new graph from excericise 4. It seems clearer and easier
to understand the most important information. I bet it depends on the situation,
though. 
```{r}
#| label: Ex-4, using facet_wrap

ggplot(midwest, aes( x= percollege, y= percbelowpoverty, color= state ) ) +
  geom_smooth( se= FALSE, show.legend= TRUE) +
labs(
    x= "Percent College Educated",
    y = "Percent Below Poverty",
    title = "Percollege vs. Percbelowpoverty")+
  facet_wrap('state')
  
```


## Exercise 5
```{r}

```


## Exercise 6
```{r}

```


## Exercise 7
```{r}

```

